---
title: "Social Media Addiction Amongst Students Dashboard"
author: "Sierra Schall and Lillian Price"
date: "today"
format:
  dashboard:
    nav-buttons: [github]
    github: https://github.com/Lillyprice/R-Dashboard-Visualization
    theme:
      - sandstone
      - theme/custom.color
    orientation: rows
    scrolling: true
    embed-resources: true
editor_options:
  chunk_output_type: console
--- 

```{r}
#| echo: false
#| output: false
#| warning: false
#| message: false

library(tidyverse)
library(kableExtra)
library(DT)
library(gt)
library(scales)
library(readr)
library(knitr)
library(paletteer)
library(shinythemes)
library(countrycode)
library(cowplot)
library(colorspace)
library(ggrepel)
library(bslib)
```


#  {.sidebar}

**Student Social Media Addiction Analysis** is built from survey-based data on social media usage habits and addiction levels among students. Each row represents one student's response.

---

**Coverage:** 110+ countries · High school, undergraduate & graduate levels

**Variables:** Age, gender, relationship status, sleep hours, most-used platform, daily usage hours, academic performance effects, and mental health & addiction scores.

---

*Dataset by Zahra Nusrat — collected to understand the extent of social media dependency among students and to identify patterns that may contribute to unhealthy usage behavior.*

```{r}
# Reading in the data 
social <- read_csv("C:/Users/lilly/GitRepo/Students Social Media Addiction.csv", show_col_types = FALSE)
social <- social |>
  group_by(Student_ID) |>
  slice(1) |>
  ungroup()
```


# Overall Effects

## Row {height="20%"}


```{r}
#| content: valuebox
#| title: Total Observations

list(
  icon  = "people-fill",
  color = "danger", 
  value = max(social$Student_ID)
)
```

```{r}
#| content: valuebox
#| title: Mean Mental Health Score 

list(
  icon  = "heart-pulse-fill",
  color = "warning", 
  value = round(mean(social$Mental_Health_Score))
)
```

```{r}
#| content: valuebox
#| title: Mean Addiction Score

list(
  icon  = "exclamation-triangle-fill",
  color = "danger", 
  value = round(mean(social$Addicted_Score))
)
```

## Row {height="40%"}

### Column {width="30%"}

```{r}
#| title: Amount of Students per Age Group
 #making a line graph of the total amount of students per age group
social |>
  count(Age) |>
  mutate( p = n/sum(n)) |>
  ggplot(aes(x = Age, y = n, group = 1)) +
  geom_line(linewidth = 2, color = "#F7A8B8") + 
  geom_point(size = 3, color = "#EE8FC6") + 
  geom_point(size = 6, color = "#F7A8B8", shape = "circle open") +
  labs(
    x = "Age Group", 
    y = "Amount"
  )


```

### Column {width="30%"}

```{r}
#| title: Average Sleep and Daily Usage by Mental Health Score
#Average Sleep and Daily Usage by Mental Health Score 
useage_summary = social |>
  mutate(MH_cate = case_when(
    Mental_Health_Score < 5 ~ "Low (0-4)", 
    Mental_Health_Score < 8 ~ "Moderate (5-7)", 
    Mental_Health_Score <= 10 ~ "High (8-10)"), 
    MH_cate = factor(MH_cate,
                     levels = c("Low (0-4)", 
                                           "Moderate (5-7)", 
                                           "High (8-10)"))) |>
  group_by(MH_cate) |>
  summarise(
    avg_sleep = mean(Sleep_Hours_Per_Night, na.rm = TRUE), 
    avg_usage = mean(Avg_Daily_Usage_Hours, na.rm = TRUE), 
    n = n()) |>
  pivot_longer(
    cols = c(avg_sleep, avg_usage), 
    names_to = "measure", 
    values_to = "average") 

ggplot(useage_summary, 
       aes(x = MH_cate, y = average, fill = measure)) +
  geom_col(position = position_dodge(width = 0.9), width = 0.8) +
  labs(
    x = "Mental Health Score", 
    y = "Average value (Hours)", 
    fill = "Measure"
  ) +  scale_fill_manual(
    values = c("#8C9A6E", "#A8BFA3"))


  
```

## Row {height="40%"}

```{r}
#| title: Table of Raw Data

#Makes our data look a lot nicer and restricts it to the first five rows
social_table = social[1:5, ]

kable(social_table, format = "html") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)
  
#Creating levels and a factor function for my bar chart
social <- social |>
  mutate(
    Academic_Level = factor(Academic_Level, 
                            levels = c("High School", "Undergraduate", "Graduate")))


```



# Students

## Row

### Column {width="50%"}

```{r}
#Students

#Creating levels and a factor function for my bar chart
social <- social |>
  mutate(
    Academic_Level = factor(Academic_Level, 
                            levels = c("High School", "Undergraduate", "Graduate")))


```

### Column {width="50%"}

```{r}
#| title: Most Used Platfrom by Academic Level

#Most Used Platform per academic level
social |>
  ggplot(aes(x = Academic_Level, y = Avg_Daily_Usage_Hours, fill = Most_Used_Platform)) + 
  geom_col(position = "fill")+ 
  labs(
    Title = "The Average Daily Usage amongst Academic Level Across the Most Used Platforms", 
    x = "Academic Level", 
    y = "Average Daily Usage (Hours)"
  ) + 
  scale_fill_paletteer_d("ggthemes::Purple_Pink_Gray") + 
  theme(legend.position = "false") + 
  coord_flip()
```

### Column {width="50%"}

```{r}
#| title: Affect of Social Media Use on Academic Performance by Grade Level
#Creating a continuous bar chart 
social |>
  count(Academic_Level, Affects_Academic_Performance) |>
  ggplot(aes(x = n, y = fct_rev(Academic_Level), 
             fill = Affects_Academic_Performance)) + 
  geom_col(position = "fill") +
  scale_fill_manual(
    values = c("#8074A8", "#C46487"), 
    guide = guide_legend(reverse = TRUE)) + 
  scale_x_continuous(labels = label_percent()) + 
  labs(x = NULL, y = NULL, fill = "Acedemic Academic Performance")
```

## Row

```{r}
#| title: How Often The Platforms are Used Amongst Respondents


#Creating a bar chart depicting how often the platforms are used amongst respondents
social |>
  count(Most_Used_Platform) |>
  mutate(p = n / sum(n)) |>
  ggplot(aes(x = reorder(Most_Used_Platform, n),
             y = n, 
             fill = Most_Used_Platform)) + 
  geom_col() + 
  scale_fill_paletteer_d("ggthemes::Purple_Pink_Gray") + 
  labs(x = "Platform", y = "Count") +
  theme_minimal() + 
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 45, hjust = 1))


```


# Countries 

## Row {height = "20%"}

```{r}
#Countries

assign_region <- function(country_vec) {
 
  df <- tibble(Country = country_vec) |>
    mutate(
      Region = countrycode(Country, origin = "country.name", destination = "region"),
      Continent = countrycode(Country, origin = "country.name", destination = "continent")
    )
  
  # Warn about unmatched countries
  unmatched <- df$Country[is.na(df$Region)]
  if (length(unmatched) > 0) {
    warning("Unmatched countries: ", paste(unique(unmatched), collapse = ", "))
  }
  
  return(df)
}

region_df <- assign_region(social$Country)

social <- social |>
  left_join(region_df, by = "Country", relationship = "many-to-many")
```


```{r}
#| content: valuebox
#| title: Total Countries

list(
  icon = "globe", 
  color = "danger", 
  value = length(unique(social$Country))
)
```

```{r}
#| content: valuebox
#| title: Total Regions

list(
  icon = "compass",
  color = "info", 
  value = length(unique(region_df$Region))
)
```

## Row {height="40%"}

### Column {width="40%"}

```{r}
#| title: Hours Slept vs Hours Spent on Social Media by Country 

country_highlight <- c(
  "Germany", "Australia", "USA", "UK", "Sri Lanka", "Russia", 
   "Maldives", "Pakistan", "Iraq", "Ghana", "Finland", "Panama", "China", 
  "Mexico", "Trinidad",  "Argentina", "Japan"
)

region_summary <- social |>
  group_by(Region, Country) |>
  summarize(
    avg_sleep = mean(Sleep_Hours_Per_Night, na.rm = TRUE),
    avg_daily_use = mean(Avg_Daily_Usage_Hours, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

region_summary <- region_summary |>
  mutate(
    label = ifelse(Country %in% country_highlight, Country, "")
  )


Region_cols <- c(
  "#F7A8B8",
  "#EE8FC6",
  "#E376D4",
  "#CC63D9",
  "#A955D4",  
  "#843FBE",  
  "#622A9A"
)

ggplot(region_summary, aes(avg_daily_use, avg_sleep)) +
  geom_point(
    aes(color = Region, fill = Region),
    size = 2.5, alpha = 0.5,
    shape = 21
  ) +
  geom_text_repel(
    aes(label = label),
    color = "black",
    size = 9/.pt,
    point.padding = 0.1, 
    box.padding = 0.6,
    min.segment.length = 0,
    max.overlaps = 1000,
    seed = 7654
  ) +
  scale_color_manual(values = darken(Region_cols, 0.3)) +
  scale_fill_manual(values = Region_cols) +
  scale_y_continuous(
    name = "Average Nightly Hours Slept",
    limits = c(4.5, 8.5),        # ← ZOOMED IN
    breaks = seq(4.5, 8.5, 0.5)
  ) +
  scale_x_continuous(
    name = "Daily Average of Hours Spent on Social Media",
    limits = c(2, 6.5),          # ← ZOOMED IN
    breaks = seq(2, 6.5, 0.5),
    expand = c(0, 0)
  ) +
  guides(
    color = guide_legend(
      nrow = 1,
      override.aes = list(
        shape = rep(21, length(unique(region_summary$Region)))
      )
    )
  ) +
  theme_minimal_hgrid(12, rel_small = 1) +
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.text = element_text(size = 9),
    legend.box.spacing = unit(0, "pt")
  )
```

### Column {width="30%"}

```{r}
#Table of the proportion of students per country
reg_social = left_join(social, region_summary)


reg_social |>
  group_by(Region) |>
  summarise(n = n_distinct(Student_ID)) |>
  mutate(p = n / sum(n)) |>
  gt() |>
  fmt_percent(
    columns = p, 
    decimals = 1
  ) |> 
  tab_style(
    style = cell_text(color = "#C46487", weight = "bold"),
    locations = cells_body(
      columns = everything(),
      rows = Region == "East Asia & Pacific"
    )
  ) |>   
  tab_style(
    style = cell_text(color = "#C46487", weight = "bold"),
    locations = cells_body(
      columns = everything(),
      rows = Region == "Europe & Central Asia"
    )
  ) |>    
  tab_style(
    style = cell_text(color = "#C46487", weight = "bold"),
    locations = cells_body(
      columns = everything(),
      rows = Region == "Latin America & Caribbean"
    )
  ) |>    
  tab_style(
    style = cell_text(color = "#C46487", weight = "bold"),
    locations = cells_body(
      columns = everything(),
      rows = Region == "Middle East & North Africa"
    )
  ) |>    
  tab_style(
    style = cell_text(color = "#C46487", weight = "bold"),
    locations = cells_body(
      columns = everything(),
      rows = Region == "North America"
    )
  ) |>    
  tab_style(
    style = cell_text(color = "#C46487", weight = "bold"),
    locations = cells_body(
      columns = everything(),
      rows = Region == "South Asia"
    )
  ) |>    
  tab_style(
    style = cell_text(color = "#C46487", weight = "bold"),
    locations = cells_body(
      columns = everything(),
      rows = Region == "Sub-Saharan Africa"
    )
  ) |> 
  cols_label(
    Region = "",
    n = "<span style='color:#B173A0'><b>Number of Students</b></span>", 
    p = "<span style='color:#B173A0'><b>Proportion of Students</b></span>", 
    .fn = md
  )
  
  
```


# Gender

```{r}
#| title: Average Mental Health Score by Gender

#Average Mental Health Score by Gender
plot_mh = social |>
  group_by(Gender) |>
  summarise(avg_mh = mean(Mental_Health_Score, na.rm = TRUE))

ggplot(plot_mh, aes( x = Gender, y = avg_mh, fill = Gender)) + 
  geom_col() + 
  labs(
       x = "Gender",
       y = "Average Mental Health Score") + 
  scale_fill_manual(
    values = c("#B173A0", "#C799BC"))
```


```{r}
#| title: "Daily Usage and Conflicts over Social Media across Gender", 

#Conflicts over social media and gender 
social |>
  ggplot(aes(Avg_Daily_Usage_Hours, as.factor(Conflicts_Over_Social_Media), fill = Gender)) +
  ggridges::geom_density_ridges(quantile_lines = TRUE, quantiles = 2, alpha = 0.7) + 
  labs(
    x = "Average Daily Usage (Hours)", 
    y = "Conflicts over Social Media", 
    fill = "Gender"
  ) +
  scale_fill_manual(
    values = c("Male" = "#E6A8A1", 
               "Female" = "#A75D74"), 
    guide = guide_legend(reverse = TRUE)) + 
  theme_minimal() 
```
