---
title: "Social Media Addiction Amongst Students Dashboard"
author: "Sierra Schall and Lillian Price"
date: "today"
format:
  dashboard:
    nav-buttons: [github]
    github: https://github.com/Lillyprice/R-Dashboard-Visualization
    theme:
      - sandstone
      - theme/custom.color
    orientation: rows
    scrolling: true
    embed-resources: true
editor_options:
  chunk_output_type: console
--- 

```{r}
#| echo: false
#| output: false
#| warning: false
#| message: false

library(tidyverse)
library(kableExtra)
library(DT)
library(gt)
library(scales)
library(readr)
library(knitr)
library(paletteer)
library(shinythemes)
library(countrycode)
library(cowplot)
library(colorspace)
library(ggrepel)
library(bslib)
```


#  {.sidebar}

Student Social Media Addiction Analysis Dataset is from survey-based information about social media usage habits and addiction levels among students. Each row is an individual student's response. It covers over 110 countries and different academic levels (high school, undergraduate, and graduate). This dataset includes demographic variables such as age, gender and relationship status, and also behavior aspects such as the average amount of sleep in hours a night, most used social media platform, and average daily use. This records the impact social media can have on sleep and if the social media usage effects them academically. This could indicate heighten feeling of anxiety, stress, and emotional dependency when not using social media. Zahra Nusrat (the owner of this particular dataset) wrote, "this dataset was collected to better understand the extent of social media dependency among students and to identify patterns that may contribute to unhealthy usage behavior". This could be useful for students, teachers, and other professionals alike to better understand the effects of social media amongst students world wide and promote healthier social media habits. 


# Overall Effects

```{r}
# Reading in the data 
social <- read_csv("~/Desktop/Stat408/R-Dashboard-Visualization/Students Social Media Addiction.csv", show_col_types = FALSE)
```

## Row {height = "20%"}

```{r}
#| content: valuebox
#| title: Total Observations

list(
  color = "danger", 
  value = max(social$Student_ID)
)
```

```{r}
#| content: valuebox
#| title: Mean Mental Health Score 

list(
  color = "warning", 
  value = mean(social$Mental_Health_Score)
)
```

```{r}
#| content: valuebox
#| title: Mean Addiction Score

list(
  color = "danger", 
  value = mean(social$Addicted_Score)
)
```

## Row {height="40%"}

### Column {width="40%"}

```{r}
#making a line graph of the total amount of students per age group
social |>
  count(Age) |>
  mutate( p = n/sum(n)) |>
  ggplot(aes(x = Age, y = n, group = 1)) +
  geom_line(linewidth = 2, color = "#F7A8B8") + 
  geom_point(size = 3, color = "#EE8FC6") + 
  geom_point(size = 6, color = "#F7A8B8", shape = "circle open") +
  labs(
    title = "Total Amount of Students per Age Group", 
    x = "Age Group", 
    y = "Amount"
  )


```

### Column {width="30%"}

```{r}
#Overall effects

#Average Sleep and Daily Usage by Mental Health Score 
useage_summary = social |>
  mutate(MH_cate = case_when(
    Mental_Health_Score < 5 ~ "Low (0-4)", 
    Mental_Health_Score < 8 ~ "Moderate (5-7)", 
    Mental_Health_Score <= 10 ~ "High (8-10)"), 
    MH_cate = factor(MH_cate,
                     levels = c("Low (0-4)", 
                                           "Moderate (5-7)", 
                                           "High (8-10)"))) |>
  group_by(MH_cate) |>
  summarise(
    avg_sleep = mean(Sleep_Hours_Per_Night, na.rm = TRUE), 
    avg_usage = mean(Avg_Daily_Usage_Hours, na.rm = TRUE), 
    n = n()) |>
  pivot_longer(
    cols = c(avg_sleep, avg_usage), 
    names_to = "measure", 
    values_to = "average") 

ggplot(useage_summary, 
       aes(x = MH_cate, y = average, fill = measure)) +
  geom_col(position = position_dodge(width = 0.9), width = 0.8) +
  labs(
    title = "Average Sleep and Daily Usage by Mental Health Score", 
    x = "Mental Health Score", 
    y = "Average value (Hours)", 
    fill = "Measure"
  ) +  scale_fill_manual(
    values = c("#8C9A6E", "#A8BFA3"))


  
```

## Row {height="40%"}

```{r}
#Makes our data look a lot nicer and restricts it to the first five rows
social_table = social[1:5, ]

kable(social_table, format = "html") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)
  
#Creating levels and a factor function for my bar chart
social <- social |>
  mutate(
    Academic_Level = factor(Academic_Level, 
                            levels = c("High School", "Undergraduate", "Graduate")))


```



# Students

### Column {width="30%"}

```{r}
#Students

#Creating levels and a factor function for my bar chart
social <- social |>
  mutate(
    Academic_Level = factor(Academic_Level, 
                            levels = c("High School", "Undergraduate", "Graduate")))

#Most Used Platform per academic level
social |>
  ggplot(aes(x = Academic_Level, y = Avg_Daily_Usage_Hours, fill = Most_Used_Platform)) + 
  geom_col(position = "fill")+ 
  labs(
    Title = "The Average Daily Usage amongst Academic Level Across the Most Used Platforms", 
    x = "Academic Level", 
    y = "Average Daily Usage (Hours)"
  ) + 
  scale_fill_paletteer_d("ggthemes::Purple_Pink_Gray")


```

### Column {width="30%"}

```{r}
#Creating a continuous bar chart 
social |>
  count(Academic_Level, Affects_Academic_Performance) |>
  ggplot(aes(x = n, y = fct_rev(Academic_Level), 
             fill = Affects_Academic_Performance)) + 
  geom_col(position = "fill") +
  scale_fill_manual(
    values = c("#8074A8", "#C46487"), 
    guide = guide_legend(reverse = TRUE)) + 
  scale_x_continuous(labels = label_percent()) + 
  labs(x = NULL, y = NULL, fill = "Acedemic Academic Performance")
```

### Column {width="30%"}

```{r}

#Creating a bar chart depicting how often the platforms are used amongst respondents
social |>
  count(Most_Used_Platform) |>
  mutate(p = n / sum(n)) |>
  ggplot(aes(x = reorder(Most_Used_Platform, n),
             y = n, 
             fill = Most_Used_Platform)) + 
  geom_col() + 
  scale_fill_paletteer_d("ggthemes::Purple_Pink_Gray") + 
  labs(x = "Platform", y = "Count") +
  theme_minimal() + 
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 45, hjust = 1))


```


# Countries 

## Row {height = "20%"}

```{r}
#Countries

assign_region <- function(country_vec) {
 
  df <- tibble(Country = country_vec) |>
    mutate(
      Region = countrycode(Country, origin = "country.name", destination = "region"),
      Continent = countrycode(Country, origin = "country.name", destination = "continent")
    )
  
  # Warn about unmatched countries
  unmatched <- df$Country[is.na(df$Region)]
  if (length(unmatched) > 0) {
    warning("Unmatched countries: ", paste(unique(unmatched), collapse = ", "))
  }
  
  return(df)
}

region_df <- assign_region(social$Country)

social <- social |>
  left_join(region_df, by = "Country", relationship = "many-to-many")
```


```{r}
#| content: valuebox
#| title: Total Countries

list(
  icon = "globe", 
  color = "danger", 
  value = length(unique(social$Country))
)
```

```{r}
#| content: valuebox
#| title: Total Regions

list(
  icon = "compass",
  color = "info", 
  value = length(unique(region_df$Region))
)
```

## Row {height="40%"}

### Column {width="40%"}

```{r}
#Countries


country_highlight <- c(
  "Germany", "Australia", "USA", "UK", "Sri Lanka", "Russia", 
   "Maldives", "Pakistan", "Iraq", "Ghana", "Finland", "Panama", "China", 
  "Mexico", "Trinidad",  "Argentina", "Japan"
)

region_summary <- social |>
  group_by(Region, Country) |>
  summarize(
    avg_sleep = mean(Sleep_Hours_Per_Night, na.rm = TRUE),
    avg_daily_use = mean(Avg_Daily_Usage_Hours, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

region_summary <- region_summary |>
  mutate(
    label = ifelse(Country %in% country_highlight, Country, "")
  )


Region_cols <- c(
  "#F7A8B8",
  "#EE8FC6",
  "#E376D4",
  "#CC63D9",
  "#A955D4",  
  "#843FBE",  
  "#622A9A"
)

ggplot(region_summary, aes(avg_daily_use, avg_sleep)) +
  geom_point(
    aes(color = Region, fill = Region),
    size = 2.5, alpha = 0.5,
    shape = 21
  ) +
  geom_text_repel(
    aes(label = label),
    color = "black",
    size = 9/.pt,
    point.padding = 0.1, 
    box.padding = 0.6,
    min.segment.length = 0,
    max.overlaps = 1000,
    seed = 7654
  ) +
  scale_color_manual(values = darken(Region_cols, 0.3)) +
  scale_fill_manual(values = Region_cols) +
  scale_y_continuous(
    name = "Average Nightly Hours Slept",
    limits = c(4.5, 8.5),        # ← ZOOMED IN
    breaks = seq(4.5, 8.5, 0.5)
  ) +
  scale_x_continuous(
    name = "Daily Average of Hours Spent on Social Media",
    limits = c(2, 6.5),          # ← ZOOMED IN
    breaks = seq(2, 6.5, 0.5),
    expand = c(0, 0)
  ) +
  guides(
    color = guide_legend(
      nrow = 1,
      override.aes = list(
        shape = rep(21, length(unique(region_summary$Region)))
      )
    )
  ) +
  theme_minimal_hgrid(12, rel_small = 1) +
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.text = element_text(size = 9),
    legend.box.spacing = unit(0, "pt")
  )
```

### Column {width="30%"}

```{r}
#Table of the proportion of students per country
reg_social = left_join(social, region_summary)


reg_social |>
  group_by(Region) |>
  summarise(n = n_distinct(Student_ID)) |>
  mutate(p = n / sum(n)) |>
  gt() |>
  fmt_percent(
    columns = p, 
    decimals = 1
  ) |> 
  tab_style(
    style = cell_text(color = "#C46487", weight = "bold"),
    locations = cells_body(
      columns = everything(),
      rows = Region == "East Asia & Pacific"
    )
  ) |>   
  tab_style(
    style = cell_text(color = "#C46487", weight = "bold"),
    locations = cells_body(
      columns = everything(),
      rows = Region == "Europe & Central Asia"
    )
  ) |>    
  tab_style(
    style = cell_text(color = "#C46487", weight = "bold"),
    locations = cells_body(
      columns = everything(),
      rows = Region == "Latin America & Caribbean"
    )
  ) |>    
  tab_style(
    style = cell_text(color = "#C46487", weight = "bold"),
    locations = cells_body(
      columns = everything(),
      rows = Region == "Middle East & North Africa"
    )
  ) |>    
  tab_style(
    style = cell_text(color = "#C46487", weight = "bold"),
    locations = cells_body(
      columns = everything(),
      rows = Region == "North America"
    )
  ) |>    
  tab_style(
    style = cell_text(color = "#C46487", weight = "bold"),
    locations = cells_body(
      columns = everything(),
      rows = Region == "South Asia"
    )
  ) |>    
  tab_style(
    style = cell_text(color = "#C46487", weight = "bold"),
    locations = cells_body(
      columns = everything(),
      rows = Region == "Sub-Saharan Africa"
    )
  ) |> 
  cols_label(
    Region = "",
    n = "<span style='color:#B173A0'><b>Number of Students</b></span>", 
    p = "<span style='color:#B173A0'><b>Proportion of Students</b></span>", 
    .fn = md
  )
  
  
```


# Gender

```{r}
#Gender


#Average Mental Health Score by Gender
plot_mh = social |>
  group_by(Gender) |>
  summarise(avg_mh = mean(Mental_Health_Score, na.rm = TRUE))

ggplot(plot_mh, aes( x = Gender, y = avg_mh, fill = Gender)) + 
  geom_col() + 
  labs(title = "Average Mental Health Score by Gender",
       x = "Gender",
       y = "Average Mental Health Score") + 
  scale_fill_manual(
    values = c("#B173A0", "#C799BC"))


#Conflicts over social media and gender 
social |>
  ggplot(aes(Avg_Daily_Usage_Hours, as.factor(Conflicts_Over_Social_Media), fill = Gender)) +
  ggridges::geom_density_ridges(quantile_lines = TRUE, quantiles = 2, alpha = 0.7) + 
  labs(
    title = "Daily Usage and Conflicts over Social Media across Gender", 
    x = "Average Daily Usage (Hours)", 
    y = "Conflicts over Social Media", 
    fill = "Gender"
  ) +
  scale_fill_manual(
    values = c("Male" = "#E6A8A1", 
               "Female" = "#A75D74"), 
    guide = guide_legend(reverse = TRUE)) + 
  theme_minimal() 



  



```
