---
title: "STAT 408 Project Proposal"
author: "Sierra Schall"
date: today 
format: html
editor: visual
---

My project team is Lilly Price and Sierra Schall, and we are planning to work with Students Social Media Addiction. They come from a vast amount of countries with enrolled students between the age of 18 to 24. It represents the struggles that a lot students face with social media and academic performance worldwide. For our project, we are going to create an R dashboard, we are going to have overall effects, age, gender, and countries as our pages. For the first page, we are going to implement some tables and graphs containing variables that relate to academic performance, mental health, sleep hours, and addiction score. The second page,we are going to look at how age relates to social media addiction. We are going to be doing something similar within the gender and countries pages.

```{r}
#| echo: false
#| output: false 

library(tidyverse)
library(kableExtra)
library(DT)
library(gt)
library(scales)
library(readr)
library(knitr)
library(paletteer)
```

```{r}
# Reading in the data 
social <- read_csv("~/Desktop/Stat408/Final Project /Students Social Media Addiction.csv", show_col_types = FALSE)

#Makes our data look a lot nicer and restricts it to the first five rows
social_table = social[1:5, ]

kable(social_table, format = "html") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)
  
#Creating levels and a factor function for my bar chart
social <- social |>
  mutate(
    Academic_Level = factor(Academic_Level, 
                            levels = c("High School", "Undergraduate", "Graduate")))

#Creating a continuous bar chart 
social |>
  count(Academic_Level, Affects_Academic_Performance) |>
  ggplot(aes(x = n, y = fct_rev(Academic_Level), 
             fill = Affects_Academic_Performance)) + 
  geom_col(position = "fill") +
  scale_fill_manual(
    values = c("#8074A8", "#C46487"), 
    guide = guide_legend(reverse = TRUE)) + 
  scale_x_continuous(labels = label_percent()) + 
  labs(x = NULL, y = NULL, fill = "Acedemic Academic Performance")

#Creating a bar chart depicting how often the platforms are used amongst respondents
social |>
  count(Most_Used_Platform) |>
  mutate(p = n / sum(n)) |>
  ggplot(aes(x = reorder(Most_Used_Platform, n),
             y = n, 
             fill = Most_Used_Platform)) + 
  geom_col() + 
  scale_fill_paletteer_d("ggthemes::Purple_Pink_Gray") + 
  labs(x = "Platform", y = "Count") +
  theme_minimal() + 
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
#Students

#Creating levels and a factor function for my bar chart
social <- social |>
  mutate(
    Academic_Level = factor(Academic_Level, 
                            levels = c("High School", "Undergraduate", "Graduate")))

#Creating a continuous bar chart 
social |>
  count(Academic_Level, Affects_Academic_Performance) |>
  ggplot(aes(x = n, y = fct_rev(Academic_Level), 
             fill = Affects_Academic_Performance)) + 
  geom_col(position = "fill") +
  scale_fill_manual(
    values = c("#8074A8", "#C46487"), 
    guide = guide_legend(reverse = TRUE)) + 
  scale_x_continuous(labels = label_percent()) + 
  labs(x = NULL, y = NULL, fill = "Acedemic Academic Performance")

#Addiction score vs performane 



```

```{r}
#Countries






```

```{r}
#Overall effects

#Average Sleep and Daily Usage by Mental Health Score 
useage_summary = social |>
  mutate(MH_cate = case_when(
    Mental_Health_Score < 5 ~ "Low (0-4)", 
    Mental_Health_Score < 8 ~ "Moderate (5-7)", 
    Mental_Health_Score <= 10 ~ "High (8-10)")) |>
  group_by(MH_cate) |>
  summarise(
    avg_sleep = mean(Sleep_Hours_Per_Night, na.rm = TRUE), 
    avg_usage = mean(Avg_Daily_Usage_Hours, na.rm = TRUE), 
    n = n()) |>
  pivot_longer(
    cols = c(avg_sleep, avg_usage), 
    names_to = "measure", 
    values_to = "average")

ggplot(useage_summary, 
       aes(x = MH_cate, y = average, fill = measure)) +
  geom_col(position = position_dodge(width = 0.9), width = 0.8) +
  labs(
    title = "Average Sleep and Daily Usage by Mental Health Score", 
    x = "Mental Health Score", 
    y = "Average value (Hours)", 
    fill = "measure"
  ) + 
  scale_fill_paletteer_d("ggthemes::Purple_Pink_Gray")
  
```

```{r}
# Age 






```
